### 2023-11-27 Munkres 1.3.tex
Day 34 HW

\documentclass{article}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage{amssymb}
\usepackage{graphicx}
\graphicspath{ {./images/} }

\title{Munkres 1.3}
\date{}
\begin{document}
\maketitle

\begin{enumerate}
    \item Define two points $(x_{0}, y_{0})$ and $(x_{1}, y_{1})$ of the plane to be equivalent if $y_{0} - x_{0}^{2} = y_{1} - x_{1}^{2}$. Check that this is an equivalence relation and describe the equivalence classes.\\
    Reflexivity: Let $(x_{0}, y_{0})$. $y_{0} - x_{0}^{2} = y_{0} - x_{0}^{2}$ so $(x_{0}, y_{0}) \sim (x_{0}, y_{0})$ for every $(x_{0}, y_{0})$.\\
    Symmetry: Let $(x_{0}, y_{0})$ and $(x_{1}, y_{1})$ such that $(x_{0}, y_{0}) \sim (x_{1}, y_{1})$. $y_{0} - x_{0}^{2} = y_{1} - x_{1}^{2}$. Equality is symmetric, so $y_{1} - x_{1}^{2} = y_{0} - x_{0}^{2}$, so $(x_{1}, y_{1}) \sim (x_{0}, y_{0})$. Thus, $(x_{0}, y_{0}) \sim (x_{1}, y_{1}) \Rightarrow (x_{1}, y_{1}) \sim (x_{0}, y_{0})$.\\
    Transitivity: Let $(x_{0}, y_{0})$, $(x_{1}, y_{1})$, and $(x_{2}, y_{2})$ such that $(x_{0}, y_{0}) \sim (x_{1}, y_{1})$ and $(x_{1}, y_{1}) \sim (x_{2}, y_{2})$. $y_{0} - x_{0}^{2} = y_{1} - x_{1}^{2}$ and $y_{1} - x_{1}^{2} = y_{2} - x_{2}^{2}$. By the transitive property of equality, $y_{0} - x_{0}^{2} = y_{2} - x_{2}^{2}$, so $(x_{0}, y_{0}) \sim (x_{2}, y_{2})$. Thus $((x_{0}, y_{0}) \sim (x_{1}, y_{1}) \land (x_{1}, y_{1}) \sim (x_{2}, y_{2})) \Rightarrow (x_{0}, y_{0}) \sim (x_{2}, y_{2})$.
    The relation is satisfies all three properties of an equivalence relation.\\

    An equivalence class of this relation determined by an element $(x_{0}, y_{0})$ is the set of all points $(x, y)$ such that $y-x^{2} = y_{0} - x_{0}^2$. In other words, let $a = y_{0} - x_{0}^2$. For any value of $x$, $(x, x^{2} + a)$ is in the equivalence class. 
    Note that for any point $(x, y)$ in the equivalence class, $(-x, y)$ is also in the equivalence class.
    \item 
\end{enumerate}
\end{document}