### 2024-01-17 CPS/LaTeX example.tex
LaTeX Example

\documentclass[10pt]{article}
\usepackage{graphicx}
\usepackage{enumerate} % to get custom labels
\usepackage{hyperref} % for footnotes to link and \hyperref
\usepackage{amsmath} % to get \text in math mode
\usepackage{amssymb} % for things like \mathcal and \mathbb
\usepackage{xcolor} % colored text
\usepackage{fancyhdr}
\usepackage[a4paper, margin=1in]{geometry}

\begin{document}
\begin{center}
    \textbf{\LaTeX{} example}
\end{center}
Hello
\begin{flushright}
    \textit{world}
    \hrule
\end{flushright}
same \hfill line

\begin{enumerate}[(1)] % why is the indentation fucked
    \item words can be:
          \begin{itemize}
              \item \footnotesize{small} (footnote size) \footnote{speaking of footnotes...}
              \item \large{big (Large)}
          \end{itemize}
    \item I can create hyperlinks to websites like \href{https://www.google.com/}{Google} or to other locations in my document (e.g., \hyperlink{a}{“more on that topic later”}).
    \item I use \texttt{array} for stuff like this:
          \begin{center}
            \renewcommand{\arraystretch}{2}
              $\begin{array}{llll} % {llll} says 4 left-aligned columns; why is vertical spacing screwy?
                      \tan{\left( - \dfrac{11 \pi}{6}\right)} & = & \tan{\dfrac{\pi}{6}}                             & \text{since} - \dfrac{11 \pi}{6} \text{ and} \dfrac{\pi}{6} \text{ are coterminal} \\[.8em]
                                                 & = & \dfrac{\sin{\dfrac{\pi}{6}}}{\cos{\dfrac{\pi}{6}}} &                                                                                  \\[.8em]
                                                 & = & \dfrac{\dfrac{1}{2}}{\dfrac{\sqrt{3}}{2}}          &                                                                                  \\[.8em]
                                                 & = & \dfrac{1}{\sqrt{3}}                              & \text{since multiplying by } \dfrac{2}{2} \text{ is legal}
                  \end{array}$
          \end{center}
    \item and \texttt{tabular} for stuff like this:
          \begin{center}
              \begin{tabular}{|l|c|r|}
                  \hline
                  Lorem & ipsum        & dolor \\
                  \S    & $\mathbb{Z}$ & \$    \\
                  \hline
              \end{tabular}
          \end{center}
    \item \textcolor{blue}{second} \quad \textcolor{red}{first}
    \item \includegraphics[scale=.1]{chrispratt.png}
\end{enumerate}
\newpage
\thispagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\hypertarget{a}{Here are those details I promised earlier.}
\end{document}