### 2024-01-22 CPS/ExpandedResume.tex
Expanded Resume

\documentclass{article}
\usepackage[a4paper, margin=0.5in]{geometry}
\usepackage{hyperref}

\begin{document}
\begin{center}
    \LARGE{\textbf{Saffron Liu}} \\
    \normalsize{}
    \href{tel:2106409659}{(210) 640-9659} \textbullet\ \href{mailto:saffron.lu.liu@gmail.com}{saffron.lu.liu@gmail.com}  \textbullet\ \href{https:\\github.com/SapphireHue}{github.com/SapphireHue}
\end{center}
\begin{flushleft}
    \Large{\textbf{Education}} \normalsize
    \rule[10pt]{\textwidth}{0.4pt}\\[-.15in]
    \textbf{Liberal Arts and Science Academy (LASA), Austin, Texas} \hfill \textsl{May 2024}\\
    \textbf{Rank}: 1/293 $\mid$ \textbf{GPA}: 4.7/4.0 $\mid$ \textbf{SAT}: 1570\\
    \textbf{Advanced Coursework}: Linear Algebra, Multivariable Calculus, Logic (Discrete Math), Computational Problem Solving, Advanced Computer Science, Digital Electronics, Latin 5\\
    \textbf{AP Courses (Scores)}: Computer Science A (5), English Language and Composition (4), Latin (5), Physics 1 (5), Physics 2 (5), United States History (5), Calculus BC (5), World History: Modern (5), English Literature and Composition (planned May 2024)

    \vspace{11pt}
    \Large{\textbf{Clubs \& Activities}} \normalsize
    \rule[10pt]{\textwidth}{0.4pt}\\[-.15in]

    \textbf{Intern @ UT Austin Matz Lab} \hfill \textsl{Aug. 2023 - Present}
    \begin{itemize}
        \setlength\itemsep{-0.2em}
        \item Assisting PhD student with projects on population genomics of coral microbes: microbes as indicators of environmental change, microbes for prediction and prevention of coral bleaching
        \item DNA extraction, library preparation (dilution, PCR, gel electrophoresis, barcoding) for next-generation sequencing
        \item Exploratory data analysis using R
    \end{itemize}

    \textbf{CyberPatriot} \hfill \textsl{Aug. 2018 - Present}
    \begin{itemize}
        \setlength\itemsep{-0.2em}
        \item National cybersecurity competition sponsored by the American Airforce Association
        \item Prepare scripts and research to secure Windows and Windows Server machines
        \item Solve forensics challenges
        \item Coordinate team roles and instruct new members annually
    \end{itemize}

    \textbf{Quiz Bowl} \hfill \textsl{Aug. 2021 - Present}
    \begin{itemize}
        \setlength\itemsep{-0.2em}
        \item Hosted intramural tournament (attended by 250+) to recruit new members
        \item Revitalized Science Bowl team
        \item Moderated 2022-2023 and 2023-2024 regional finals for Science Bee, Geography Bee, and History Bee
        \item Facilitated communication between Longhorn and LASA Quiz Bowl to host 2023 Prometheus Science Bowl
        \item 5th place at local tournament, qualifying for the 2024 High School National Championship Tournament
    \end{itemize}

    \textbf{Latin Club}\\
    Praetor (academic manager; elected position) \hfill \textsl{May 2023 - Present}
    \begin{itemize}
        \setlength\itemsep{-0.2em}
        \item Run twice weekly certamen (Latin Jeopardy!) practices and provide players with study resources
        \item Manage club tournament/convention registration
        \item Organized and hosted 2nd annual LASA Certamen tournament (20+ teams from Austin, San Antonio, and Houston)
    \end{itemize}
    Co-President (elected position) \hfill \textsl{Aug. 2022 - May 2023}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Planned and ran weekly club and certamen (Latin Jeopardy) meetings with consistent membership
        \item Managed club tournament/convention registration and convention logistics (travel, lodgings, meals)
        \item Helped host 1st annual LASA Certamen tournament
        \item Promoted the Latin class/club at showcase and outreach events, increasing enrollment and membership (~15/yr)
        \item Brought back annual Broomball fundraiser (\$600) after it stopped due to COVID-19, allowing purchases for the Latin classroom and subsidizing club convention fees
        \item Worked with teacher sponsor to reorganize and more clearly define club roles
    \end{itemize}
    Scribe (elected position) \hfill \textsl{Aug. 2021 - May 2022}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Compiled and distributed officer meeting notes
        \item Helped host 2-day long area convention, overcoming challenges posed by new/unfinished building, new teacher sponsor, and first in-person convention after COVID-19
    \end{itemize}
    Certamen Player / Latin Club Member \hfill \textsl{Jan. 2021 - Present}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Study and mentor younger certamen players in Greek Mythology and Latin Literature year-round
        \item Compete in certamen and academic testing about classical subjects from city to national level
    \end{itemize}

    \newpage
    \textbf{Fundamentals of Computer Science: Teaching Assistant} \hfill \textsl{Aug. 2023 - Present}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Work one-on-one with students to clarify topics and debug code
        \item Answer questions and assist in grading for a class of ~30
    \end{itemize}

    \textbf{LASA Literary Club: Co-founder}\hfill \textsl{Aug. 2020 - May 2022}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Created club to combat social isolation due to COVID
        \item Curated book list and created book groups based on member preferences
        \item Led small discussion groups
    \end{itemize}

    \textbf{Austin Public Library Summer Volunteer} \hfill \textsl{Jun. 2022 - Aug. 2022}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Set up 3 different gaming consoles for the Girl’s Gaming Club
        \item Oversaw group of ~30 people for 4 hours at a time
        \item Commended by program supervisor for independently continuing the program while they were out sick
    \end{itemize}

    \textbf{Piano} \hfill \textsl{2011 - 2020 \& Aug. 2023 - Present}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Enrolled in private lessons from 2011 to 2020; currently enrolled in school independent study piano class
        \item Recent repetoire: \textit{Preludio} and \textit{Melodioso} from \textit{8 Valses Poeticos}, Granados; \textit{Sarabande} and \textit{Gigue} from English Suite No. 2 in A minor, BWV 807, J.S. Bach; \textit{Presto alla tedesca} from Sonata in G, Op. 79, Beethoven; \textit{Nocturne} in G minor, Op. 37 No. 1, Chopin; \textit{Alla turca} from Piano Sonata No. 11, Mozart
    \end{itemize}

    \textbf{Web Development}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Developed \href{https://github.com/ks4729/ks4729.github.io/blob/main/.github/workflows/Spotify_Content.yml}{script} to automatically retrieve user's top songs with Spotify API and update website with Github Actions
        \item Created \href{https://github.com/SapphireHue/Certamen-Buzzer}{website} to parse certamen questions using regular expressions and incrementally display text to users, helping increase access to study materials and combat high barrier to entry
        \item Created \href{https://www.geogebra.org/m/pmvspthz}{GeoGebra file} to create visualization for multivariable calculus class using JavaScript
        \item Projects in progress: database of questions using MongoDB; multiplayer buzzer system using Node.js and Express.js
    \end{itemize}

    \textbf{Other Clubs/Organizations}: Programming in Practice, Women+ in Computer Science, National Honor Society
    \textbf{Other Competitions}: USACO, picoCTF, CyberStart America, Science Bowl, UIL Computer Science, NACLO
    \textbf{Additional Activities}: \href{https://photos.app.goo.gl/tEm2jn1kG2grJVUn6}{crochet}, \href{https://photos.app.goo.gl/ZvdBS4s2AQgb9bm28}{origami}

    \vspace{11pt}
    \Large{\textbf{Education}} \normalsize
    \rule[10pt]{\textwidth}{0.4pt}\\[-.15in]
    \textbf{Academic Awards}\\
    LASA Class of 2024 Valedictorian (2022, 2023)\\
    National Merit Semifinalist (2023)\\
    Rochester Institute of Technology Women in STEM High School Award (2023)\\
    Trustee Award (top 10\% of class) (2023, 2022)\\
    AP Scholar with Distinction (2023)

    \vspace{11pt}
    \textbf{Computer Science}\\
    CyberPatriot: Platinum National Semifinalist (top 7.5\% nationally) (2023)\\
    National Cybersecurity Scholarship Foundation (through CyberStart): Scholar (2023), Finalist (2022)\\
    UIL Computer Science: 2nd place written test and 1st place team at Regionals (2023), 2nd place team at District (2023)\\
    USA Computing Olympiad: Silver Division (2023)\\
    picoCTF: top 10\% (2023)\\

    \vspace{11pt}
    \textbf{Classics}\\
    Nationals 2023: 1st place advanced open certamen, 2nd place Greek mythology, 8th place Latin literature, 6th place academic heptathlon (includes Greco-Roman mythology, Roman history, Greek, and Latin)\\
    State 2023: 1st place advanced competitive certamen, 1st place Greek mythology, 1st place Latin literature\\
    State 2022: 1st place advanced open certamen, 2nd place Greek mythology, 1st place Latin literature\\
    State 2021: 3rd place intermediate competitive certamen, 10th place Greek mythology\\
    Area 2023: 1st place advanced competitive certamen, 1st place Greek mythology, 1st place Latin literature\\
    Area 2022: 2nd place advanced competitive certamen, 1st place Greek mythology, 1st place Latin literature\\
    Area 2021: 1st place intermediate competitive certamen, 2nd place Greek mythology\\

    \vspace{11pt}
    \textbf{Other}\\
    ABRSM Piano Performance Grade 8 (2020)\\
    ABRSM Music Theory Grade 5, passed with merit (2017)\\

\end{flushleft}

\end{document}