### 2024-01-23 CPS/tikz example.tex
Tikz Example - got it to work by reinstalling latex 🥲

\documentclass[11pt]{amsart}
\usepackage[margin=1in]{geometry}            
\usepackage[parfill]{parskip}

% \usepackage{ifplatform, pdftexcmds, catchfile}
% \usepackage{ifplatform, pdftexcmds, catchfile}

\input{Bailey_Barnes_Preamble.sty}

\begin{document}
\begin{center}\textbf{Tikz examples}

\begin{tikzpicture}[scale=1]
    \draw[dashed, gray] (-2,-2)node[below left]{$(-2,-2)$} grid (4,2)node[above right]{$(4,2)$};
    \draw[ultra thick, <->](-1,-1)--(3,1)node[right]{$A$};
    \draw[ultra thick] (-1,1) circle (1);
    \draw[ultra thick, domain=-2:3] plot (\x, {sin(\x r)});
    \draw[ultra thick, domain=0:4, red, <->] plot (\x, {-(pow(\x,2)-4*\x+2)});
\end{tikzpicture}

\begin{tikzpicture}[scale=.5]
    \draw[ultra thick, red, domain=0:1800, samples=500] plot(\x: {\x / 360});
    \draw[ultra thick, blue, domain=0:180,samples=500, fill=blue!15] plot (\x: {6*sin(5*\x)});
    \foreach \x in {1, 2, ..., 6}
        \draw[gray] (0,0) circle (\x);

    \foreach \a in {15, 30, ..., 345}
        \draw[gray] (0:0) -- (\a:6);
    \draw[<->] (-7,0)--(7,0) node[right]{$x$};
    \draw[<->] (0,-7)--(0,7) node[above]{$y$};

\end{tikzpicture}
\end{center}
\end{document}